Open Data, Open Source and Open Standards in chemistry - The Blue Obelisk five years on

### 2011-04-26 paper.tex
Added text on Open Data

A considerable stumbling block in advocating the release of scientific
data as Open Data has been how exactly to define Open. A major step
forward was the launch in 2010 of the Panton Principles for Open Data
in Science \cite{WebPanton}. This formalises the idea that Open Data maximises the
possibility of reuse and repurposing, the fundamental basis
of how science works. These principles recommend that published data
be licensed explicitly, and preferably under CC0 (Creative Commons `No
Rights Reserved') \cite{WebCC0}. This license allows others to use the data for any
purpose whatsoever without any barriers.

@@

      \end{tabular}
      }
  \subsection*{Table 2 - Sample table title}
    Large tables are attached as separate files but should
    still be described here.

>>>>

A considerable stumbling block in advocating the release of scientific
data as Open Data has been how exactly to define Open. A major step
forward was the launch in 2010 of the Panton Principles for Open Data
in Science \cite{WebPanton}. This formalises the idea that Open Data maximises the
possibility of reuse and repurposing, the fundamental basis
of how science works. These principles recommend that published data
be licensed explicitly, and preferably under CC0 (Creative Commons `No
Rights Reserved', also known as CCZero) \cite{WebCC0}. This license allows others to use the 
data for any purpose whatsoever without any barriers. Other licenses
compatible with the Panton Principles include the
Open Data Commons Public Domain Dedication and Licence (PDDL), the
Open Data Commons Attribution License, and the
Open Data Commons Open Database License (ODbL) (see
http://www.opendefinition.org/licenses/\#Data).

Despite this positive news, little chemical data has become
available from the traditional chemical fields of organic,
inorganic, solid state chemistry. Table~2 lists a few notable
exceptions: ChemPedia (a now discontinued crowd-sourcing project),
CrystalEye (http://wwmm.ch.cam.ac.uk/crystaleye/),
and the Open Notebook Science Solubility
data~\cite{ONS2010}. There is also data available using licenses
not compatible with the Panton Principles, but where the user
is allows to modify and redistribute the data. A new data
set in this category is the data from the ChEMBL database, 
which is available under the Creative Commons Share-Alike
Attribution license~\cite{Overington2009}.

Importantly, publishing data as CC0 is become easier now that
website are becoming available to simplify publishing data. Two
projects that can be mentioned in this context are FigShare
(http://figshare.com/), where the data behind unpublished figures
can be hosted, and Dryad (http://datadryad.org/) where data
behind publications can be hosted. Initiatives like this make
it possible to host small amounts of data, and those combinded
are expected to become soon a substantial knowledge base.

@@

      \end{tabular}
      }
  \subsection*{Table 2 - Open Data in chemistry.}
    Overview of major open chemical data available under a license or waiver
    compatible with the Panton Principles.
    \par \mbox{}
    \par
    \mbox{
\begin{tabular}{|l|l|l|}
  %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
  \hline Name & License/Waiver & Description \\ \hline
  ChemPedia & CC0 & Crowd-sourced chemical names. Project discontinued. \\ \hline
  CrystalEye     & PPDL & Crystal structures from primary literature. \\ \hline
  ONS Solubility & CC0 & Solubility data for various solvents. \\ \hline
\end{tabular}
      }
